Discrete Quantum Theories and Computing

### 2018-12-12 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

In conventional quantum theory and computation, the probability amplitudes
are modeled by the continuum of complex numbers. This kind of model
doesn't take into account that human doesn't have the ability to distinguish
quantities differing only by an arbitrarily small extent. One possibility
is that the probability amplitudes in quantum theories are quantized.
This leads us to consider quantum theories and computing over finite
fields. Even if the probability amplitudes are actually continuous,
since human has no ability to distinguish the probability computed
by these probability amplitudes, it may be desirable to incorporate
the idea of measurement precision into the quantum theory, and consider
quantum interval-valued probability.
\begin{sidewaystable}
\caption{Comparison among our models, where the upper half of the table lists
their quantum mechanical definitions and properties, and the lower
half of the table lists their computational power.}

>>>>

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, modeling
quantum mechanics and computation over the continuum of complex numbers
may not capture our ability of predicting quantum phenomenon and the
power of quantum computation. Being agnostic about whether the reality
is ultimately discrete or continuous, we considered two types of models
to incorporate the limitation of distinguishability. By assuming the
quantum states are quantized and distinguishable, instead of the conventional
theory over the continuum of complex numbers, we considered the quantum
theories and computing over finite fields. Shifting our attention
from the reality to the limitation of distinguishability itself, the
second model incorporate the finite precision measurement into the
quantum theory called quantum interval-valued probability measures.

Even before quantum mechanics is discovered, human tried to characterize
the physical quantities and controlled the quantities and operations
to simulate and predict unknown quantities. Discrete quantities can
be easily characterized by whole numbers~\cite{ParkerBaldridge2004}.
Even if the proportion of whole numbers are fractions but not whole
numbers, the operation among fractions can be considered as the operation
among whole numbers after we find the common denominator, and the
resulting numbers faithfully represent the quantities they represent.
In general, discrete computation can be reliably repeated with exact
equivalence, and any digital computer can faithfully simulate a Turing
machine limited only by the available memory and time~\cite{Turing_1937}.

The non-apparently discrete quantities and their operations confused
many cultures from the very beginning. An ancient Chinese philosopher
Chuang-tzu (庄子) has expressed ``If a one-foot-long stick is cut into
halves every day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
\begin{sidewaystable}
\caption{Comparison among our models, where the upper half of the table lists
their quantum mechanical definitions and properties, and the lower
half of the table lists their computational power.}
